Revisiting Short-Chain Fatty Acids and the Microbiota in Colorectal Cancer

### 2018-05-23 submission/manuscript.Rmd
add references to intro

Colorectal cancer (CRC) is currently the third leading cancer-related cause of death within the US []. The envrionment has been attributed as an important risk factor for CRC []. This has lead to the investigation of how the microbiota may exacerbate or cause tumorigensis [] and whether the bacterial community is altered []. Multiple reports in case/control studies on the microbiota and CRC have identified taxa commonly associated with the normal colon to be decreased in patients with carcinoma tumors []. Many of these taxa actively produce short-chain fatty acids (SCFAs) from fiber that is part of our general diet []. The most extensively studied of these SCFAs are acetate, butyrate, and propionate []. Overall, CRC is a growing health problem and the microbiota through SCFAs have shown promise as a viable treatment option.

Within model systems SCFAs, in particular butyrate, have shown positive results in CRC treatment []. In particular, butyrate has been shown to inhibit cancer cell growth in *in vitro* systems []. Additionally, supplementation with food sources that bacteria use to create these SCFAs may also be able to confer beneficial effects. Fiber supplementation in mouse models of CRC caused an overall reduction in tumor burden while increasing SCFA concentrations []. Although these model systems provide important preliminary evidence towards the ability of SCFAs to reduce and treat tumors, the studies reporting benefit in humans has been less convincing.

There is a lack of evidence on the benefit of increasing SCFA concentrations to protect against CRC in human populations. The initial case/control studies that investigated SCFA concentrations in CRC have found that patients with carcinomas had lower concentrations of acetate, butyrate, and propionate versus patients with adenomas and individuals with normal colons []. Although this would argue that increasing SCFA concentrations could be protective against tumorigenesis, fiber supplementation in randomized controlled trials have consistently failed to protect against tumor recurrence []. These findings would argue against the utility of treatments that aim to use SCFAs to reduce or protect against tumorigenesis. Given the lack of clear evidence in human studies of the benefit of SCFAs in CRC there is a need for more investigation into this area.

>>>>

Colorectal cancer (CRC) is currently the third leading cancer-related cause of death within the US [@Haggar2009; @Siegel2016]. The envrionment has been attributed as an important risk factor for CRC [@Lichtenstein2000]. This has lead to the investigation of how the microbiota may exacerbate or cause tumorigensis [@Zackular2013; @Baxter2014; @Zackular2015] and whether the bacterial community is altered [@Shah2017; @meta_analysis_crc_Sze2018]. Multiple reports in case/control studies on the microbiota and CRC have identified taxa commonly associated with the normal colon to be decreased in patients with carcinoma tumors [@Zeller2014; @Baxter2016; @Kostic2011]. Many of these taxa actively produce short-chain fatty acids (SCFAs) from fiber that is part of our general diet [@Holscher2017]. The most extensively studied of these SCFAs are acetate, butyrate, and propionate [@Louis2016]. Overall, CRC is a growing health problem and the microbiota through SCFAs have shown promise as a viable treatment option.

Within model systems SCFAs, in particular butyrate, have shown positive results in CRC treatment [@test_OKeefe2016]. In particular, butyrate has been shown to inhibit cancer cell growth in *in vitro* systems [@Encarnao2018]. Additionally, supplementation with food sources that bacteria use to create these SCFAs may also be able to confer beneficial effects. Fiber supplementation in mouse models of CRC caused an overall reduction in tumor burden while also increasing SCFA concentrations [@Bishehsari2018]. Although these model systems provide important preliminary evidence towards the ability of SCFAs to reduce and treat tumors, the studies reporting benefit in humans has been less convincing.

There is a lack of evidence on the benefit of increasing SCFA concentrations to protect against CRC in human populations. The initial case/control studies that investigated SCFA concentrations in CRC have found that patients with carcinomas had lower concentrations of acetate, butyrate, and propionate versus patients with adenomas and individuals with normal colons [@Ohigashi2013]. Although this would argue that increasing SCFA concentrations could be protective against tumorigenesis, fiber supplementation in randomized controlled trials have consistently failed to protect against tumor recurrence [@Yao2017]. These findings would argue against the utility of treatments that aim to use SCFAs to reduce or protect against tumorigenesis. Given the lack of clear evidence in human studies of the benefit of SCFAs in CRC there is a need for more investigation into this area.
